---
title: "What Make Coffee Popular?"
author: "Ken Trinh, Michelle Lee, Tanmay Mahapatra"
date: "03/30/2022"
output:
  bookdown::pdf_document2: 
    toc: true
    number_sections: true
    toc_depth: 3
---

\newpage
\setcounter{page}{1}
```{r load packages and set options, include=FALSE}
#install.packages('patchwork')
#install.packages('moments')
install.packages("tidytuesdayR")
library(tidyverse) 
library(magrittr)
library(knitr)
library(patchwork)
library(moments)

theme_set(theme_bw())

options(tinytex.verbose = TRUE)
knitr::opts_chunk$set(echo=FALSE, message=FALSE)
```


```{r load-data, message=FALSE, inlcude=FALSE, echo=FALSE}
tuesdata <- tidytuesdayR::tt_load('2020-07-07')
tuesdata <- tidytuesdayR::tt_load(2020, week = 28)
coffee_ratings <- tuesdata$coffee_ratings
```


# Introduction
-- Work in a causal relationship based research design


Coffee is one of the most popular non-alcoholic beverages globally prized for its aroma and caffeine content. Coffee is brewed from roasted beans of the plant species Coffea, which is native to sub-Saharan Africa and individual islands in the Indian Ocean. Ever since its discovery, coffee has grown to more than 70 tropical countries, making up a multi-billion dollar market. Despite being consumed for over 120 years, there are still new trends entering the market every year, shifting the popularity of different coffee types. Some of these trends include different beans, different caffeine concentration, and different brewing techniques. Because of the constant changes in coffee trends, everything from the variety of the plant, the chemistry of the soil, the weather, the amount of rainfall and sunshine, and even the precise altitude at which the coffee grows can affect the taste of the final product. 

In an effort to understand what drives the popularity of coffee, our team proposed an investigation following the question:

\begin{quote}
  \textit{How does the country of origin affect the popularity of a coffee?}
\end{quote}

-- remove processing method and include flavor
Prior to looking at the data set, we conceptualized the popularity of a coffee as a rating point, out of 100, given to a cup of coffee and the country of origin as the country that the coffee bean came from. Some covariates that we planned to consider in the study are the different `processing_method`. The `processing_method` is conceptualized as the methods that the coffee bean is processed by such as Natural/Dry or Washed/Wet.  

Our analysis will use the 2020 coffee data set gather by tidytuesday\footnote{Rfordatascience. (2020). Tidytuesday/readme.md at master · rfordatascience/tidytuesday. 
GitHub. Retrieved March 23, 2022, from 
https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-07-  
07/readme.md}. We planned to use some of the observational data collected as follows:

* First, we use column named `total_cup_points` that came with the data set as our measurement for our conceptualization for the popularity of coffee. This column is already given as a numerical value out of 100, so the closer to 100 means the more popular that specific cup of coffee.

* -------Work in what model we planned to us here.


After the data is collected and processed, we plan to tackle this problem by building several models to capture the question. The first model will capture the effect of country of origin on coffee ratings. Because there could be other confounding variables, we want to also look into other covariates including flavor and processing method. These features will be added to a second model to capture their effect. Additional models will be needed in the study to provide a more conclusive result.

Our research team believes that answering this question would enable us to identify how different factors affect the rating of coffee. Ultimately, the result of our analysis would help ACME with identifying new coffee products to include in the store by prioritizing features as well as with improving our marketing strategy.


